---
title: "AirBnB Destination Prediction"
author: "Sharlene, Sonali, Bhushan"
date: "Tuesday, March 08, 2016"
output:
  ioslides_presentation:
    logo: airbnb.png
    smaller: yes
    widescreen: yes
  beamer_presentation: default
  slidy_presentation: default
---

##AirBnB Booking Prediction on Kaggle

Training data consisting of around 214000 users.

- NaiveBayes
- Xgboost(Gradient Boosting)
- Exploratory analysis

##Various Packages Used
- xgboost
- lubridate
- e1071
- rworldmap
- reshape
- sqldf

```{r, echo=FALSE}
library(xgboost)
library(sqldf)
library(reshape)
library(rpart.plot)
library(stringr)
library(caret)
library(car)
library(e1071)
library(sqldf)
library(dplyr)
library(MASS)
library(ggmap)
library(rworldmap)
library(ggplot2)
library(lubridate)
library(rpart)
library(caTools)
library(randomForest)
set.seed(1)
```


```{r,echo=FALSE}
setwd("E:/Datasets/Airbnb")
age_buckets_data=read.csv("age_gender_bkts.csv")
df_train_Data = read.csv("train_users_2.csv")
copyofRaw = df_train_Data
labels = copyofRaw['country_destination']
countryDetails = read.csv("countries.csv")
```

##Some Information about Data

```{r}
names(copyofRaw)
```


##Destination Data

```{r,echo=FALSE}
countries= table(df_train_Data$country_destination)
barplot(countries*100/sum(countries), xlab = "Destination", ylab = "Percentage")


```


##Types of Booking

```{r,echo=FALSE}
tempDF = df_train_Data
tempDF = tempDF %>% mutate(country_destination=
                                ifelse(country_destination== 'AU' |
                                      country_destination== 'CA'|
                                      country_destination== 'FR'|
                                      country_destination== 'GB'|
                                      country_destination== 'ES'|
                                        country_destination== 'IT'|
                                        country_destination== 'PT'|
                                        country_destination== 'NL'|
                                        country_destination== 'DE'|
                                        country_destination== 'other'
                                        ,'International',
                                      ifelse(country_destination== 'US',
                                             'Domestic',
                                             'No booking')))

international = table(tempDF$country_destination)
barplot(international*100/sum(international), xlab = "Type of Travel", ylab = "Percentage",col="#FD5C64")

```


##Gender Data

```{r,echo=FALSE}
genders = table(df_train_Data$gender)
barplot(genders*100/sum(genders), xlab = "Gender", ylab = "Percentage",col="#FD5C64")

```

##Age Data
```{r,echo=FALSE}
qplot(x=df_train_Data$age,geom="histogram",
      binwidth = 10,  
      main = "Histogram for Age", 
      xlab = "Age",  
      fill=I("blue"), 
      col=I("red"), 
      alpha=I(.2))




```


##Age Data - Post cleanup
```{r,echo=FALSE}
dfAge = df_train_Data
dfAge$age = as.integer(dfAge$age)
dfAge = dfAge %>% mutate(age = 
    ifelse(dfAge$age>1920 & dfAge$age<2000,2016-dfAge$age,
           ifelse(dfAge$age>16 & dfAge$age<96,dfAge$age,NaN))
  )

qplot(x=dfAge$age,geom="histogram",
      binwidth = 10,  
      main = "Histogram for Age", 
      xlab = "Age",  
      fill=I("blue"), 
      col=I("red"), 
      alpha=I(.2))

```


##User accounts and bookings over time

```{r,echo=FALSE}
series = df_train_Data
series$date_account_created = as.Date(series$date_account_created)
series$date_first_booking = as.Date(as.character(series$date_first_booking), format="%Y-%m-%d")

series$week_created = as.Date(cut(series$date_account_created,
                                     breaks = "week", 
                                     start.on.monday = FALSE))
series$week_booked = as.Date(cut(series$date_first_booking,
                                    breaks = "week", 
                                    start.on.monday = FALSE))

weekly_booked = group_by(series, date= week_booked) %>%
    summarise(booking_count = n())

weekly_created = group_by(series, date = week_created) %>%
    summarise(created_count = n())

weekly_booked = weekly_booked[!is.na(weekly_booked$date),]
# head(weekly_booked)


weekly_stats = merge(weekly_booked, weekly_created, by="date")

weekly_stats = melt(weekly_stats, id.vars = "date")

ggplot(weekly_stats, aes(date, value, col = variable)) +
    geom_line()

```

##Weekday corresponding to maximum booking

```{r,echo=FALSE}
dfDay = df_train_Data %>% filter(date_first_booking != '' & country_destination != 'NDF')
dfDay$date_first_booking = factor(dfDay$date_first_booking)

dfDay$day = weekdays(as.Date(dfDay$date_first_booking))

dfDay$day <- factor(dfDay$day, levels= c("Sunday", "Monday", 
    "Tuesday", "Wednesday", "Thursday", "Friday", "Saturday"))
count = table(dfDay$day)
barplot(count, xlab = "Weekday", ylab = "Total Bookings",col="#FD5C64")


```


##Cleaning Data

- Cleaning up age variables
- Creating new variables such as ageProvded, difference of days in activity and booking
- Month of the booking
- Separating factors into their levels

```{r,echo=FALSE}
cleanDF = df_train_Data
cleanDF = cleanDF %>% mutate(ageProvided = ifelse(is.na(age),0,1))
cleanDF = cleanDF %>% mutate(age = 
                           ifelse(age>1916 & age<2016,2016-age,
                                  ifelse(age>16 & age<100,age,NaN))
)
cleanDF[is.na(cleanDF)] = -1
#cleanDF = na.omit(cleanDF)
# table(cleanDF$age)
cleanDF = cleanDF %>% mutate(bookingDone = ifelse(cleanDF$country_destination=='NDF'
                                                  ,0,1))

cleanDF$gender=as.factor(cleanDF$gender)
cleanDF$age=as.integer(cleanDF$age)
cleanDF$signup_method=as.factor(cleanDF$signup_method)
cleanDF$signup_flow=as.factor(cleanDF$signup_flow)
cleanDF$language=as.factor(cleanDF$language)
cleanDF$affiliate_provider=as.factor(cleanDF$affiliate_provider)
cleanDF$first_affiliate_tracked=as.factor(cleanDF$first_affiliate_tracked)
cleanDF$signup_app=as.factor(cleanDF$signup_app)
cleanDF$first_device_type=as.factor(cleanDF$first_device_type)
cleanDF$first_browser=as.factor(cleanDF$first_browser)
cleanDF$country_destination=as.factor(cleanDF$country_destination)
cleanDF$bookingDone = as.factor(cleanDF$bookingDone)
cleanDF$ageProvided = as.factor(cleanDF$ageProvided)
cleanDF$id=as.factor(cleanDF$id)
cleanDF$date_account_created=as.factor(cleanDF$date_account_created)
cleanDF$timestamp_first_active=as.factor(cleanDF$timestamp_first_active)
cleanDF$date_first_booking=as.factor(cleanDF$date_first_booking)

```


##Predicting Successful Booking
Now we will try to break this problem into small parts. We will now try to predict whether a user books AirBnB or not.

```{r,echo=FALSE}
dataset = cleanDF
# for (i in 1:ncol(dataset)) {
#   #  if (sum(as.numeric(is.na(df_train_Data[, i]))) > 0) {
#   num_missing = sum(as.numeric(is.na(dataset[, i])))
#   print(paste0(colnames(dataset)[i], ":  ", num_missing))
#   
#   # }
# }



set.seed(121)
splDS = sample.split(dataset$bookingDone, SplitRatio = 0.7)
trainDS = subset(dataset, splDS==TRUE)
testDS= subset(dataset, splDS==FALSE)

actualBooking = testDS$bookingDone
testDS = subset(testDS,select = -c(bookingDone))

modelDS = naiveBayes(trainDS$bookingDone ~ gender+age+signup_method+
                        language+signup_app+affiliate_channel+affiliate_provider+
                        first_affiliate_tracked+
                        first_device_type+first_browser, data = trainDS)

predDS=predict(modelDS, testDS)
# predDS
confNBDS = table(predDS, actualBooking)
confNBDS
chisq.test(confNBDS)
```

##Recursive Tree for Booking prediction
```{r,echo=FALSE}
Tree = rpart(trainDS$bookingDone ~ gender+age+signup_method+
               language+signup_app+affiliate_channel+affiliate_provider+
                 first_affiliate_tracked+ageProvided+
                 first_device_type+first_browser, data = trainDS, method="class", minbucket=25)
# # 
PredictCART = predict(Tree, newdata = testDS, type = "class")
confRpart = table(PredictCART, actualBooking)
confRpart
prp(Tree,type=1,extra=2)
```


##Predicting type of Booking

```{r,echo=FALSE}
bookedData = dataset%>% filter(bookingDone==1)
diff = as.Date(as.character(bookedData$date_first_booking), format="%Y-%m-%d")-
  as.Date(as.character(bookedData$date_account_created), format="%Y-%m-%d")

# table(diff)
bookedData = bookedData %>% mutate(difference = ifelse(diff> 0 ,diff,-1))

bookedData = bookedData %>% mutate(typeofbooking=
                                      ifelse(country_destination== 'US'
                                             ,'Domestic',
                                             'International'
                                      ))
# table(bookedData$typeofbooking)
# names(bookedData)
bookedData$date_first_booking = factor(bookedData$date_first_booking)

dfb = as.data.frame(str_split_fixed(bookedData$date_first_booking, '-', 3))
bookedData['dfb_month'] = as.factor(dfb[,2])

bookedData$typeofbooking = as.factor(bookedData$typeofbooking)

set.seed(121)
splType = sample.split(bookedData$bookingDone, SplitRatio = 0.7)
trainType = subset(bookedData, splType==TRUE)
testType= subset(bookedData, splType==FALSE)
bookingType = testType$typeofbooking
testType = subset(testType,select = -c(typeofbooking))

modelType = naiveBayes(typeofbooking ~ gender+age+signup_method+
                          language+signup_app+affiliate_channel+affiliate_provider+
                          first_affiliate_tracked+dfb_month+
                          first_device_type+first_browser+ageProvided+difference+date_first_booking, data = trainType)

predType=predict(modelType, testType)
confType = table(predType,bookingType)

confType
chisq.test(confType)
```



##Predicting Destination country

NaiveBayes
```{r,echo=FALSE}
interBooking = bookedData %>% filter(typeofbooking=='International')
set.seed(121)
interBooking$country_destination = factor(interBooking$country_destination)

splCoun = sample.split(interBooking$country_destination, SplitRatio = 0.7)
trainCoun = subset(interBooking, splCoun==TRUE)
testCoun= subset(interBooking, splCoun==FALSE)

countries = testCoun$country_destination
testCoun = subset(testCoun,select = -c(country_destination))

modelCoun = naiveBayes(trainCoun$country_destination ~ ., data = trainCoun)
predCoun=predict(modelCoun, testCoun)
confCoun = table(predCoun,countries)

print(confCoun)

```

##Predicting Destination country

```{r,echo=FALSE}
chisq.test(confCoun)
sum = 0
for(i in 1:nrow(confCoun)){
  sum = sum + confCoun[i,i]
}
print(sum/sum(confCoun))
```

##Predicting Destination country
Now we sample 500 rows for each destination and check output. 

```{r,echo=FALSE}
dother=sqldf("select * from interBooking where country_destination=='other' LIMIT 500") 
dAU=sqldf("select * from interBooking where country_destination=='AU' LIMIT 500") 
dCA=sqldf("select * from interBooking where country_destination=='CA' LIMIT 500") 
dDE=sqldf("select * from interBooking where country_destination=='DE' LIMIT 500") 
dES=sqldf("select * from interBooking where country_destination=='ES' LIMIT 500") 
dFR=sqldf("select * from interBooking where country_destination=='FR' LIMIT 500") 
dGB=sqldf("select * from interBooking where country_destination=='GB' LIMIT 500") 
dIT=sqldf("select * from interBooking where country_destination=='IT' LIMIT 500")
dNL=sqldf("select * from interBooking where country_destination=='NL' LIMIT 500") 
dPT=sqldf("select * from interBooking where country_destination=='PT' LIMIT 217") 

dALL=sqldf("select * from dother union select * from dAU union select * from dCA union select * from dDE union select * from dES union select * from dFR union select * from dGB union select * from dIT union select * from dNL union select * from dPT")

splCoun1 = sample.split(dALL$country_destination, SplitRatio = 0.7)
trainCoun1 = subset(dALL, splCoun1==TRUE)
testCoun1= subset(dALL, splCoun1==FALSE)

countries = testCoun1$country_destination
testCoun1 = subset(testCoun1,select = -c(country_destination))

modelCoun1 = naiveBayes(trainCoun1$country_destination ~ ., data = trainCoun1)
predCoun1=predict(modelCoun1, testCoun1)
confCoun1 = table(predCoun1,countries)

confCoun1

```

##Predicting Destination country
```{r,echo=FALSE}
chisq.test(confCoun1)
sum = 0
for(i in 1:nrow(confCoun1)){
  sum = sum + confCoun1[i,i]
}
print(sum/sum(confCoun1))

```


##Gradient Boosting
- eXtreme Gradient Boosing
- Produces a prediction model in the form of an ensemble of weak prediction models
- **Bagging**: Build models on random samples of data
- **Boosting**: More weight given to samples which are hard to classify rather than sample

xgb = xgboost(**data** = data.matrix(train_X), 
               **label** = train_Y, 
               **eta** = 0.1,
               **max_depth** = 9, 
               **nround**=25, 
               **subsample** = 0.5,
               **colsample_bytree** = 0.5,
               **seed** = 1,
               **eval_metric** = "merror",
               **objective** = "multi:softprob",
               **num_class** = 12,
               **nthread** = 3
)

##Important Parameters
- **Objective**: Type of model that needs to be trained. In our case we have used multi:softprob to do multi -class classification
- Output contains predicted probability of each data point belonging to each class
- **Depth**: Maximum Depth of trees(Default=6)
- **nround**: Number of passes on the data
- **nthread**: Number of cpu threads being used
- **eta**: Shrinks the feature weights after boosting to avoid overfitting
- **subsample**: Subsample Ratio to training instances

##Predicting Destination country

```{r,echo=FALSE}
dfXGB = cleanDF#rbind(df_train22,df_test22)

# split date_account_created in year, month and day
dac = as.data.frame(str_split_fixed(dfXGB$date_account_created, '-', 3))
dfXGB['dac_year'] = dac[,1]
dfXGB['dac_month'] = dac[,2]
dfXGB['dac_day'] = dac[,3]
dfXGB = dfXGB[,-c(which(colnames(dfXGB) %in% c('date_account_created')))]

# split timestamp_first_active in year, month and day
dfXGB['tfa_year'] = as.factor(substring(as.character(dfXGB[,'timestamp_first_active']), 1, 4))
dfXGB['tfa_month'] = as.factor(substring(as.character(dfXGB['timestamp_first_active']), 5, 6))
dfXGB['tfa_day'] = as.factor(substring(as.character(dfXGB['timestamp_first_active']), 7, 8))
dfXGB = dfXGB[,-c(which(colnames(dfXGB) %in% c('timestamp_first_active')))]

# We split the factor variables in multiple columns and then run the xgboost
features = c('gender', 'signup_method', 'signup_flow', 'language', 'affiliate_channel', 'affiliate_provider', 'first_affiliate_tracked', 'signup_app', 'first_device_type', 'first_browser')
dummies = dummyVars(~ gender + signup_method + signup_flow + language + affiliate_channel + affiliate_provider + first_affiliate_tracked + signup_app + first_device_type + first_browser, data = dfXGB)
dfXGB_ohe = as.data.frame(predict(dummies, newdata = dfXGB))

dfXGB_combined = cbind(dfXGB[,-c(which(colnames(dfXGB) %in% features))],dfXGB_ohe)

set.seed(121)
spl = sample.split(dfXGB_combined$country_destination, SplitRatio = 0.7)
train_sample = subset(dfXGB_combined, spl==TRUE)
test_sample = subset(dfXGB_combined, spl==FALSE)

train_X = train_sample[,c(1:3,5:144)]
train_Y = train_sample["country_destination"]

#train22_X["country_destination"]

test_X = test_sample[,c(1:3,5:144)] 
test_Y = test_sample["country_destination"]
ADest = test_Y

# str(ADest)
# str(test_sample)

train_Y = recode(train_Y$country_destination,"'NDF'=0; 'US'=1; 'other'=2; 'FR'=3; 'CA'=4; 'GB'=5; 'ES'=6; 'IT'=7; 'PT'=8; 'NL'=9; 'DE'=10; 'AU'=11")
train_Y = as.numeric(levels(train_Y))[train_Y]

test_Y = recode(test_Y$country_destination,"'NDF'=0; 'US'=1; 'other'=2; 'FR'=3; 'CA'=4; 'GB'=5; 'ES'=6; 'IT'=7; 'PT'=8; 'NL'=9; 'DE'=10; 'AU'=11")
test_Y = as.numeric(levels(test_Y))[test_Y]

X_test = test_X

xgb = xgboost(data = data.matrix(train_X), 
               label = train_Y, 
               eta = 0.1,
               max_depth = 9, 
               nround=25, 
               subsample = 0.5,
               colsample_bytree = 0.5,
               seed = 1,
               eval_metric = "merror",
               objective = "multi:softprob",
               num_class = 12,
               nthread = 3
)

y_pred = predict(xgb, data.matrix(X_test))

predictions = as.data.frame(matrix(y_pred, nrow=12))
rownames(predictions) = c('NDF','US','other','FR','CA','GB','ES','IT','PT','NL','DE','AU')
# topPredictions_3 = as.data.frame(apply(predictions, 2, function(x) names(sort(x)[12:10])))


topPredictions_3_F = as.vector(apply(predictions, 2, function(x) names(sort(x)[12:10])))

ids = NULL
actualDest = NULL
for (i in 1:NROW(X_test)) {
  idx = X_test$id[i]
  ids = append(ids, rep(idx,3))
}

Predict = NULL
Predict$id = ids
Predict$P_country = topPredictions_3_F

Predict = as.data.frame(Predict)

top1 = as.vector(apply(predictions, 2, function(x) names(sort(x)[12])))

top1 = recode(top1,"'NDF'=0; 'US'=1; 'other'=2; 'FR'=3; 'CA'=4; 'GB'=5; 'ES'=6; 'IT'=7; 'PT'=8; 'NL'=9; 'DE'=10; 'AU'=11")
#top1 = as.numeric(levels(top1))[top1]
length(top1)

```

##Prediction Format
```{r,echo=FALSE}
head(Predict)
t = table(top1,test_Y)
t
print((t[1,1]+t[2,2])/sum(t))
```

##Does Signup_App have a relationship with Destination?
```{r,echo=FALSE}
t = table(df_train_Data$signup_app, df_train_Data$country_destination) 
df_App = as.data.frame(t)
names(df_App) = c("signup_app", "destination_country", "Count")
ggplot(df_App, aes(signup_app, Count, fill=destination_country)) + geom_bar(stat="identity")
```


##Does Signup_App have a relationship with Destination?
```{r,echo=FALSE}
t = table(df_train_Data$signup_app, df_train_Data$country_destination) 
df_App = as.data.frame(t / rowSums(t))
names(df_App) = c("signup_app", "destination_country", "Ratio")
ggplot(df_App, aes(signup_app, Ratio, fill=destination_country)) + geom_bar(stat="identity")

```

##What effect does Language have upon Destination Country?
```{r,echo=FALSE}
t = table(df_train_Data$language, df_train_Data$country_destination) 
df_Lang = as.data.frame(t / rowSums(t))
names(df_Lang) = c("Language", "destination_country", "Ratio")
ggplot(df_Lang, aes(Language, Ratio, fill=destination_country)) + geom_bar(stat="identity")
```

##What effect does Signup_method have upon Destination Country?
```{r,echo=FALSE}
t = table(df_train_Data$signup_method, df_train_Data$country_destination) 
df_Method = as.data.frame(t)
names(df_Method) = c("Signup_Method", "destination_country", "Ratio")
ggplot(df_Method, aes(Signup_Method, Ratio, fill=destination_country)) + geom_bar(stat="identity")
```

##What effect does Signup_method have upon Destination Country?
```{r,echo=FALSE}
t = table(df_train_Data$signup_method, df_train_Data$country_destination) 
df_Method = as.data.frame(t / rowSums(t))
names(df_Method) = c("Signup_Method", "destination_country", "Ratio")
ggplot(df_Method, aes(Signup_Method, Ratio, fill=destination_country)) + geom_bar(stat="identity")
```

##Future scope for AirBnB
```{r,echo=FALSE}
# newmap <- get_map()
# 
# plot(newmap)
# points(countryDetails$lng_destination,countryDetails$lat_destination, col = "red")

countryCount = df_train_Data %>% group_by(country_destination) %>% summarise(count = n())
countryCount = subset(countryCount , country_destination!='NDF')
countryCount = subset(countryCount , country_destination!='other')

finalCountry = merge(countryDetails, countryCount, by = 'country_destination')

mp <- NULL
mapWorld <- borders("world", colour="gray50", fill="gray50") # create a layer of borders
mp <- ggplot() +   mapWorld

#Now Layer the cities on top
mp <- mp+ geom_point(data = finalCountry, aes(x = lng_destination, y = lat_destination, size=count), color="red") 
mp
```

##What could be done better?
- Feature engineering 
- Lots of scope for improving accuracy of NaiveBayes
- Implementing other prediction algorithms

```{r,echo=FALSE}
#Attemp to count number of predictions correct in top3
# top3 = t(topPredictions_3)
# head(top3)
# 
# top3 = as.data.frame(top3)
# colnames(top3) = c("P1","P2","P3")
# 
# top3$P1 = recode(top3$P1,"'NDF'=0; 'US'=1; 'other'=2; 'FR'=3; 'CA'=4; 'GB'=5; 'ES'=6; 'IT'=7; 'PT'=8; 'NL'=9; 'DE'=10; 'AU'=11")
# top3$P2 = recode(top3$P2,"'NDF'=0; 'US'=1; 'other'=2; 'FR'=3; 'CA'=4; 'GB'=5; 'ES'=6; 'IT'=7; 'PT'=8; 'NL'=9; 'DE'=10; 'AU'=11")
# top3$P3 = recode(top3$P3,"'NDF'=0; 'US'=1; 'other'=2; 'FR'=3; 'CA'=4; 'GB'=5; 'ES'=6; 'IT'=7; 'PT'=8; 'NL'=9; 'DE'=10; 'AU'=11")
# 
# top3$P1 = as.numeric(levels(top3$P1))[top3$P1]
# top3$P2 = as.numeric(levels(top3$P2))[top3$P2]
# top3$P3 = as.numeric(levels(top3$P3))[top3$P3]
# top3$count = 0
# head(top3)
# 
# correct_Classified = 0
# check = function(D, e) {
#   for(i in 1:nrow(D)){
#     for(j in 1:ncol(D)-1){
#       
#     if(e[i] %in% D[i,j]) {
#         D[i]$count = 1
#     }
#   }
# }
# }
# correct = check(top3,test_Y)


```

